\fred{A Unified Treatment of Tractability of Approximation Problems}

### 2023-07-06 GeneralizedTractability_230607.tex
Updates from Overleaf

\documentclass[sort&compress]{elsarticle}
\usepackage{latexsym,amsfonts,amsmath,graphics,mathtools,amssymb,booktabs,xspace}
\usepackage{epsfig}
\usepackage[notref,notcite]{showkeys}
%\usepackage[active]{srcltx}
\usepackage{fdsymbol}
\usepackage{enumitem}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage[normalem]{ulem}
\usepackage{algorithmic}
\usepackage{hyperref}
\usepackage{cleveref}

@@

%certain important notations for key symbols
\newcommand{\thed}{\clubsuit}
\newcommand{\theM}{\varheartsuit}
\newcommand{theUB}{\spadesuit}

@@

\title{\fred{A Unified Treatment of Tractabilty of Approximation Problems}}

\author[1]{Fred J. Hickernell}
\ead{hickernell@iit.edu}
\author[2]{Peter Kritzer}
\ead{peter.kritzer@oeaw.ac.at}
\author[1]{Onyekachi Osisiogu}
\date{\today}
\affiliation[1]{organization={Department of Applied Mathematics, Illinois Institute of Technology},
addressline={10 West 32${}^{\text{nd}}$ Street, Room 220},
postcode={60616},
city={Chicago, IL},
country={United States}}
\affiliation[2]{organization={Johann Radon Institute for Computational and Applied Mathematics (RICAM), Austrian Academy of Sciences},
addressline={Altenbergerstr. 69},
postcode={4040},
city={Linz},
country={Austria}}

@@

TODO: We need to check the relation of our work to that of Michael Gnewuch and Henryk. We do not need tensor product structure: do they need that assumption? The summation conditions on the eigenvalues might be a significant difference between our results and theirs.

@@

@@

For this case where arbitrary linear functionals are allowed, the optimal solution operator is
\[
\APP_d(f,\varepsilon) = \sum_{i=1}^n \peter{\SOL_d(u_{i,d})} \ip[\cf_d]{f}{u_i},
\]
and the information complexity of our linear problem is
\begin{equation}\label{eq:comp_def}
\comp(\varepsilon, d) = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\}.
\end{equation}

The question of tractability is one of determining how fast the \peter{information} complexity increases as $\varepsilon$ tends to zero and/or as $d$ tends to infinity.  We want to bound the complexity in terms of a simple function, $T$, of $\varepsilon^{-1}$ and $d$ and identify simple conditions on the singular values that are equivalent to this bound holding. \peter{[PETER: Maybe this would be a good place to refer to the earlier work of Michael and Henryk? I will take care of it... .]}

In this article, we investigate a \emph{generalized tractability function}, denoted by $T$:
\begin{equation} \label{eq:Tspec}
    T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow (0,\infty).
\end{equation}
The conditions we require $T$ to satisfy are described in  Section \ref{sec:not}, however, the basic idea is that we define our approximation problem to be tractable if 
$\comp(\varepsilon, d) \le  C_{\vp}\, T(\varepsilon^{-1},d,\vp)$ for some constant, $C_{\vp}$, depending only on the parameter $\vp$.  In practice, we expect $T(\varepsilon^{-1},d,\vp)$ to increase sub-exponentially  with increasing $\varepsilon^{-1}$ and/or $d$, however most of our theorems do not require such an assumption.

\Cref{thm_main_strong_tract2,thm_main_tract2,thm_main_rest_strong_tract,thm_main_rest_tract} all take the form

\peter{[PETER: Not meaning to sound too negative, but I am not sure I like the notation with the symbols below. Is it possible to use something different instead?]}
\fred{Fred: I have now given them names above in lines 18-20, so we can experiment with different symbols or letters.}
\begin{multline}
  \exists C_{\vp} \text{ with }
    \comp(\varepsilon, d) \le  C_{\vp}\, T(\varepsilon^{-1},\thed,\vp) \quad \forall (\varepsilon^{-1},d) \in \Omega \\
    \iff   \exists L_{\vp} \text{ with }
    \sup_{d \in \naturals}
     \sum_{i = \lceil L_{\vp}\, T(0,\thed,\vp) \rceil}^{\theUB(d)} \frac{1}{T(\lambda^{-1}_{i,d},\thed,\vp)}< \infty, \\
     \thed \in \{1,d\}, \ \theM(d) := \inf \{\varepsilon : (\varepsilon^{-1},d) \in \Omega \}, \ 
     \theUB(d) := \begin{cases}
        \min\{n \in \natzero : \lambda_{n+1,d} \le \theM(d) \}, &   \theM(d) > 0, \\
        \infty, & \theM(d) = 0.
        \end{cases}
\end{multline}
\Cref{thm:subhT} is of similar, but not equivalent, form.

\peter{TODO: Maybe add a short paragraph on the notation used in the paper and on its structure? In particular, we should introduce
an ordering of the parameters $\vp$.}\\
%\kachi{Throughout the article, we use the following notation.}\\

@@

\subsection{Notation and fundamental definitions} \label{sec:not}
Let $T$ be a function given in \eqref{eq:Tspec}.
A tractability function $T$ is a function of simple form that provides an upper bound on the \peter{information} complexity of a problem.  Some examples of tractability functions commonly considered are provided in Table \ref{tab:commonT}.

\begin{table} 
    \caption{Common forms of the tractability function, $T$. \peter{[PETER: I changed the notation below, using ``algebraic" and ``exponential" first, to be in line with the literature.]}}
{\small
\begin{equation*}
	\begin{array}{r@{\quad}c@{\quad}cc}
		\text{Tractability type} & T(\varepsilon^{-1},d,\vp)
		& T(0,d,\vp)\\
		\toprule
		\text{\peter{algebraic strong polynomial}} & \max(1,\varepsilon^{-p}) & 1\\
		\text{\peter{algebraic polynomial}} & \max(1,\varepsilon^{-p})d^{q} & d^{q}\\
		\text{\peter{exponential strong polynomial}} &  [\max(1,\log(1 + \varepsilon^{-1})]^p & 1\\
		\text{\peter{exponential polynomial}} &
		[\max(1,\log(1 + \varepsilon^{-1}))]^p  d^{q} & d^{q} \\
        \text{\peter{algebraic quasi-polynomial}} & 
        \exp\{p(1+\log(\max(1,\varepsilon^{-1}))(1+\log(d))\} &
        \exp\{p(1+\log(d))\}\\
        \text{\peter{exponential quasi-polynomial}} & 
        \exp\{p(1+\log(\max\{1,\log(1+\varepsilon^{-1})\}))(1+\log(d))\} &
        \exp\{p(1+\log(d))\}
	\end{array}
\end{equation*}}
\label{tab:commonT}
\end{table}

@@

\section{Examples} \label{sec:examples}
%In this section, we illustrate Theorems \ref{thm_main_strong_tract2} and \ref{thm_main_tract2} for the types of tractability given in the table in Section \ref{sec:not}.
In this section, we present six examples relating to Theorem \ref{thm_main_strong_tract2} and Theorem \ref{thm_main_tract2}, both of which were introduced in Section \ref{sec:spt}, for the various types of tractability that are listed in Table \ref{tab:commonT}. In subsequent sections, we will continue to present various examples.

@@

\begin{example}[\peter{Exponential strong} polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= \left[\max\{1,\log(1+\varepsilon^{-1})\}\right]^{\peter{\vp}}, \quad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty),
\]
%\kachi{Alternatively:
%\[
%T(\varepsilon^{-1},d,\vp) = (1+\log\max\{1,\varepsilon^{-1}\})^p
%\]}
where \peter{$\vp=p$} is a scalar. Then Theorem \ref{thm_main_strong_tract2} yields the equivalent condition for \peter{exponential strong polynomial tractability:} 
\[
 \sup_{d\in\naturals} \sum_{i=L_{p} }^\infty (\min\{1,\log(1+\lambda_{i,d}\})^{p}  < \infty,\quad  \text{ for some } p \ge 0, L_{p} \in \naturals.
\] In this case the optimal $p$ is the infimum of all $p$ for which the above condition holds.
\end{example}
\begin{example}[Exponential polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)=\left[\max\{1,\log(1+\varepsilon^{-1}\})\right]^p\, d^q, \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2, 
\] where $\vp = (p,q)$. Then Theorem \ref{thm_main_tract2} yields  the equivalent condition for exponential polynomial tractability
\[
 \sup_{d\in\naturals}\, d^{-q} \sum_{i=\lceil L_{(p,q)} d^{q}\rceil}^\infty [\min\{1,\log(1+\lambda_{i,d}\})]^p < \infty  \text{ for some } p,q \ge 0, L_{(p,q)} \in \naturals.
\]
\end{example}

\peter{[PETER: Are we sure that the conditions in Examples 3 and 4 are correct? And, moreover, these conditions should recover the corresponding conditions in \cite{KriWoz19a}. The conditions look different at first sight, but probably it can be shown that they are equivalent... .}

In the following example, we provide a brief summary of these notions of tractability. We would like to emphasize that this example was previously known (see \cite{}). 
\begin{example}[(Strong) Separable tractability]
Suppose that the tractability function is defined by
\[
 T(\varepsilon^{-1},d,\vp)= \phi(\varepsilon^{-1},p)\, \psi(d,q)
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2.
\]
with non-decreasing functions $\phi : (0,\infty) \times (0,\infty) \to (0,\infty)$ and $\psi : \naturals  \times (0,\infty) \to (0,\infty)$ and where $\vp = (p,q)$, hence $T$ is non-decreasing in all variables.

We say that a problem is defined as $T$-separable tractable with parameter $\vp$ iff there exists a positive constant $C_{\vp}$ which does not depend on $\varepsilon$ and $d$ such that
\[
\comp(\varepsilon,d) \leq  C_{\vp} \,  \phi(\varepsilon^{-1},p)\, \psi(d,q)\qquad \forall \varepsilon >0, \, d\in\naturals.
\]
A problem is strongly $T$-separable tractable with parameter $\vp$ iff the complexity is independent of $d$, that is, there exists a positive constant $C_{\vp}$ which again does not depend on $\varepsilon$ and $d$ such that
\[
\comp(\varepsilon,d) \leq C_{\vp}\,\phi(\varepsilon^{-1},p)\qquad \forall \varepsilon >0, \, d\in\naturals,
\] where in this case $\vp = p$ is a scalar. 
Furthermore, we assume the following as in \eqref{eq:Tlim}, \eqref{eq:T0}, and \eqref{eq:ptauassume}. We assume that

\[
\lim_{\varepsilon\rightarrow0}\phi(\varepsilon^{-1},p) = \infty \qquad \forall d\in\naturals, p >0.
\] since there are an infinite number of positive eigenvalues. Furthermore, we have that $\phi(\cdot,p)$ is non-decreasing hence we define the following limit
\[ \lim_{\varepsilon \rightarrow \infty}\phi(\varepsilon^{-1},p)\geq 1.
\] 
Additionally, we have the following technical assumptions. Suppose that, for any real $\tau_1, \tau_2>0$ and any $p,q>0$, we have that there exists $\hat{K}_{p,\tau_1}$ and $\hat{K}_{q\tau_2}$ such that
\begin{align*}
     [\phi(\varepsilon^{-1},p)]^{\tau_1} &\le \hat{K}_{p,\tau_1} \phi (\varepsilon^{-1},\tau_1 p)\\
      [\psi(d,q)]^{\tau_2} &\le \hat{K}_{q,\tau_2} \psi (d,\tau_2 q) \qquad \forall \, \varepsilon \in (0,\infty),\, d\in \naturals.
\end{align*} where $\hat{K}_{p,\tau_1}>0$ is a constant that may depend on $p$ and $\tau_1$, but is independent of $\varepsilon$, and where $\hat{K}_{q,\tau_2}>0$ is a constant that may depend on $q$ and $\tau_2$, but is independent of $d$.

Thus, Theorem \ref{thm_main_strong_tract2} yields the following equivalent condition for strongly $T$-separable tractability
\begin{equation*} \label{eq:strong_tractiff}
     \sup_{d \in \naturals} \sum_{i = L_p}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} < \infty,
\end{equation*}
for some $p>0$ and $L_p \in \naturals$. Furthermore, the optimal $p$ is also infimum of all $p$.
Theorem \ref{thm_main_tract2} yields the following equivalent condition for 
separable tractability
\[
 \sup_{d \in \naturals} \frac{1}{\psi(d,q)} \sum_{i = \lceil L_{(p,q)}\psi(d,q)\rceil}^\infty \frac{1}{\phi(\lambda_{i, d}^{-1},p)} < \infty,
\] for some $p,q >0$ and $L_{(p,q)}\in\naturals$.
\end{example}
Next, we will introduce another summarized notion of tractability. 
\begin{example}[Non-separable tractability]
    Suppose we define the tractability function $T$ by
\[
 T(\varepsilon^{-1},d,p)= \exp\{p \, \phi(\varepsilon^{-1})\, \psi(d)\} = [\exp\{ \phi(\varepsilon^{-1})\}]^{p \, \psi(d)}
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, p\in [0,\infty).
\]  with non-decreasing functions $\phi : (0,\infty) \to (0,\infty)$ and $\psi : \naturals \to (0,\infty)$. 
We say that a problem is non-separable tractable with parameter $p$ iff there exists a positive $C_{p}$ such that

\[\comp(\varepsilon,d) \leq  C_{p}\exp\{p(\phi(\varepsilon^{-1})\psi(d))\}.\]

Hence we can apply Theorem \ref{thm_main_tract2} as follows. We have the following equivalent condition for non-separable tractability
\[
%\sup_{d\in\mathbb{N}}\frac{1}{\psi(d)}\sum_{i=Lp\,\psi(d)} \phi(\lambda_{i,d})\quad \text{for some } p \geq 0, L_p \in \naturals.
\sup_{d\in\mathbb{N}}\sum_{i=L_p\,\phi(0)^{p \, \psi(d)}} [\exp\{ \phi(\lambda_{i,d}^{-1})\}]^{p \, \psi(d)}\quad \text{for some } p \geq 0, L_p \in \naturals.
\]
\end{example}
For $\phi$ and $\psi$ of special form, the condition above can be simplified further as in the following two examples of non-separable tractability. 
\begin{example}[Algebraic quasi-polynomial tractability]

As defined in \cite{GneWoz11a}, we say that a problem is \emph{quasi-polynomially tractable (QPT)} if there exist two constants $C$, $p > 0$, such that
\begin{equation}\label{expqpt}
    \comp(\varepsilon,d) \leq C\exp\{p(1+\log(\varepsilon^{-1}))(1+\log(d))\}
\end{equation}
for all $\varepsilon \in (0,1]$ and $d\in\mathbb{N}$.

Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,p)= \exp\{p(1+\log(\max\{1,\varepsilon^{-1}\}))(1+\log (d))\}  \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, p \in [0,\infty).
\] This can be rewritten as 
\[
T(\varepsilon^{-1},d,p) = e^p[\max\{1,\varepsilon^{-1}\}]^{p(1+\log d)}d^p
\]
Observe that the function $T$ is a function of $\max\{1,\varepsilon^{-1}\}$ and $d$ which are, polynomial or not exponential in powers of $\max\{1,\varepsilon^{-1}\}$ and $d$. Hence we can apply Theorem \ref{thm_main_tract2} as follows. We have the following equivalent condition for quasi-polynomial tractability in the algebraic case
\[
\sup_{d\in\mathbb{N}}\,d^{-p}\sum_{i=Lp\,d^{p}} \left(\min\{1,\lambda_{i,d}\} \right)^{p(1+\ln d)} < \infty \qquad \text{for some } p \geq 0, L_p \in \naturals.\]
\end{example}
Next, we will consider the exponential case as follows.
\begin{example}[Exponential quasi-polynomial tractability]

>>>>

\documentclass[sort&compress]{elsarticle}
\usepackage{latexsym,amsfonts,amsmath,graphics,mathtools,amssymb,booktabs,xspace}
\usepackage{epsfig}
\usepackage[notref,notcite]{showkeys}
%\usepackage[active]{srcltx}
\usepackage{fdsymbol}
\usepackage{enumitem}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage[normalem]{ulem}
\usepackage{algorithmic}
\usepackage{hyperref}
\usepackage{mathrsfs}
\usepackage{cleveref}

@@

%certain important notations for key symbols
%\newcommand{\thed}{\clubsuit}
%\newcommand{\theM}{\varheartsuit}
%\newcommand{\theUB}{\spadesuit}
\newcommand{\thed}{\delta}
\newcommand{\theM}{\mathscr{E}}
\newcommand{\theUB}{\mathscr{U}}

@@

\title{\fred{A Unified Treatment of Tractability of Approximation Problems}}

\author[1]{Fred J. Hickernell}
\ead{hickernell@iit.edu}
\author[2]{Peter Kritzer}
\ead{peter.kritzer@oeaw.ac.at}
\author[1]{Onyekachi Osisiogu}\ead{oosisiogu@iit.edu}
\date{\today}
\affiliation[1]{organization={Department of Applied Mathematics, Illinois Institute of Technology},
addressline={10 West 32${}^{\text{nd}}$ Street, Room 220},
postcode={60616},
city={Chicago, IL},
country={United States}}
\affiliation[2]{organization={Johann Radon Institute for Computational and Applied Mathematics (RICAM), Austrian Academy of Sciences},
addressline={Altenbergerstr. 69},
postcode={4040},
city={Linz},
country={Austria}}

@@

@@

TODO: We should check consistency of notation when using $\max$ or $\min$. Sometimes we use ``$(\ldots)$'' for the argument, sometimes we use ``$\{\ldots\}$''

@@

For this case where arbitrary linear functionals are allowed, the optimal solution operator is
\[
\APP_d(f,\varepsilon) = \sum_{i=1}^n \peter{\SOL_d(u_{i,d})} \ip[\cf_d]{f}{u_i},
\]
and the information complexity of our linear problem is
\begin{equation}\label{eq:comp_def}
\comp(\varepsilon, d) = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\}.
\end{equation}
\peter{
Strictly speaking, we are considering the \textit{absolute} error criterion in this paper. 
Alternatively, one could also consider the \textit{normalized} error criterion, where the error in \eqref{eq:errcrit} is divided by the so-called \textit{initial error}, which is the error without sampling the function, or, equivalently, the operator norm of $\SOL_d$. In the latter case, it is known (see, e.g., \cite{NovWoz08a}) that the information complexity is given by 
\[
\comp^{\rm norm}(\varepsilon, d) = \min \{n \in \natzero : \lambda_{n+1, d} \le \varepsilon\, \lambda_{1, d}\}.
\]
This means that, essentially, in the tractability analysis below, all criteria would be normalized by the first eigenvalue $\lambda_{1,d}$.

For the sake of brevity and clarity of notation, we restrict ourselves to the absolute error criterion here, but presumably, analogous results hold for the normalized setting. 
}

The question of tractability is one of determining how fast the \peter{information} complexity increases as $\varepsilon$ tends to zero and/or as $d$ tends to infinity.  We want to bound the complexity in terms of a simple function, $T$, of $\varepsilon^{-1}$ and $d$ and identify simple conditions on the singular values that are equivalent to this bound holding. \peter{We would like to point out that generalized tractability specified by function $T$ has been considered before in the literature. In particular, there are several papers by Gnewuch and Wo\'{z}niakowski (see \cite{GW06,GW08,GW09,GneWoz11a} ), and also \cite[Chapter 8]{NovWoz08a} is devoted to this topic. To our best knowledge, however, these references mainly focus on the special case tensor product problems, which we do not do here, and the conditions for tractability derived there are mostly in terms of properties of the function $T$, and not in terms of sums of (functions of) the eigenvalues of a problem. Hence, we consider the present paper an addition to what is presently known on this subject.}

In this article, we investigate a \emph{generalized tractability function}, denoted by $T$:
\begin{equation} \label{eq:Tspec}
    T :(0,\infty) \times \mathbb{N} \times [0,\infty)^s \rightarrow (0,\infty).
\end{equation}
The conditions we require $T$ to satisfy are described in  Section \ref{sec:not}, however, the basic idea is that we define our approximation problem to be tractable if $\comp(\varepsilon, d) \le  C_{\vp}\, T(\varepsilon^{-1},d,\vp)$ for some constant, $C_{\vp}$, depending only on the parameter $\vp$.  In practice, we expect $T(\varepsilon^{-1},d,\vp)$ to increase sub-exponentially  with increasing $\varepsilon^{-1}$ and/or $d$, however, most of our theorems do not require such an assumption.

\Cref{thm_main_strong_tract2,thm_main_tract2,thm_main_rest_strong_tract,thm_main_rest_tract} all take the form
\begin{multline}
  \exists \,C_{\vp} > 0 \text{ with }
    \comp(\varepsilon, d) \le  C_{\vp}\, T(\varepsilon^{-1},\thed,\vp) \quad \forall (\varepsilon^{-1},d) \in \Omega \\
    \iff   \exists L_{\vp} \text{ with }
    \sup_{d \in \naturals}
     \sum_{i = \lceil L_{\vp}\, T(0,\thed,\vp) \rceil}^{\theUB(d)} \frac{1}{T(\lambda^{-1}_{i,d},\thed,\vp)}< \infty, \\
     \thed \in \{1,d\}, \ \theM(d) := \inf \{\varepsilon : (\varepsilon^{-1},d) \in \Omega \}, \ 
     \theUB(d) := \begin{cases}
        \min\{n \in \natzero : \lambda_{n+1,d} \le \theM(d) \}, &   \theM(d) > 0, \\
        \infty, & \theM(d) = 0.
        \end{cases}
\end{multline}
\Cref{thm:subhT} is of similar, but not equivalent, form.

\peter{TODO: Maybe add a short paragraph on the notation used in the paper and on its structure? In particular, we should introduce
an ordering of the parameters $\vp$.}\\
\\

@@

\subsection{Notation and fundamental definitions} \label{sec:not}
Let $T$ be a function given in \eqref{eq:Tspec}.
A tractability function $T$ is a function of a simple form that provides an upper bound on the \peter{information} complexity of a problem.  Some examples of tractability functions commonly considered are provided in Table \ref{tab:commonT}.
\kachi{}
\begin{table} 
    \caption{Common forms of the tractability function, $T$. \peter{[PETER: I changed the notation below, using ``algebraic" and ``exponential" first, to be in line with the literature.]}}
{\small
\begin{equation*}
	\begin{array}{r@{\quad}c@{\quad}cc}
		\text{Tractability type} & T(\varepsilon^{-1},d,\vp)
		& T(0,d,\vp)\\
		\toprule
		\text{\peter{algebraic strong polynomial}} & \max(1,\varepsilon^{-p}) & 1\\
		\text{\peter{algebraic polynomial}} & \max(1,\varepsilon^{-p})d^{q} & d^{q}\\
		\text{\peter{exponential strong polynomial}} &  [\max(1,\log(1 + \varepsilon^{-1})]^p & 1\\
		\text{\peter{exponential polynomial}} &
		[\max(1,\log(1 + \varepsilon^{-1}))]^p  d^{q} & d^{q} \\
        \text{\peter{algebraic quasi-polynomial}} & 
        \exp\{p(1+\log(\max(1,\varepsilon^{-1}))(1+\log(d))\} &
        \exp\{p(1+\log(d))\}\\
        \text{\peter{exponential quasi-polynomial}} & 
        \exp\{p(1+\log(\max\{1,\log(1+\varepsilon^{-1})\}))(1+\log(d))\} &
        \exp\{p(1+\log(d))\}
	\end{array}
\end{equation*}}
\label{tab:commonT}
\end{table}

@@

\section{Examples} \label{sec:examples}
%In this section, we illustrate Theorems \ref{thm_main_strong_tract2} and \ref{thm_main_tract2} for the types of tractability given in the table in Section \ref{sec:not}.
In this section, we present six examples relating to Theorem \ref{thm_main_strong_tract2} and Theorem \ref{thm_main_tract2}, both of which were introduced in Section \ref{sec:spt}, for the various notions of tractability that are listed in Table \ref{tab:commonT}. In subsequent sections, we will continue to present various examples.

@@

\begin{example}[\peter{Exponential strong} polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)= [\max\{1,\log(1+\varepsilon^{-1})\}]^{\peter{\vp}}, \quad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty),
\]

where \peter{$\vp=p$} is a scalar. Then Theorem \ref{thm_main_strong_tract2} yields the equivalent condition for \peter{exponential strong polynomial tractability:} 
\begin{equation}\label{eq:exp_spt}
 \sup_{d\in\naturals} \sum_{i=L_{p} }^\infty \frac{1}{[\max\{1,\log(1+\lambda_{i,d}^{-1})\}]^{p}}  < \infty,\quad  \text{ for some } p \ge 0, L_{p} \in \naturals.
\end{equation}
In this case the optimal $p$ is the infimum of all $p$ for which the above condition holds.
\end{example}
\begin{example}[Exponential polynomial tractability]
Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,\vp)=\left[\max\{1,\log(1+\varepsilon^{-1})\}\right]^p\, d^q, \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2, 
\] where $\vp = (p,q)$. Then Theorem \ref{thm_main_tract2} yields  the equivalent condition for exponential polynomial tractability:
\begin{equation}\label{eq:exp_pt}
 \sup_{d\in\naturals}\, d^{-q} \sum_{i=\lceil L_{(p,q)} d^{q}\rceil}^\infty \frac{1}{[\max\{1,\log(1+\lambda_{i,d}^{-1})\}]^p} < \infty  \text{ for some } p,q \ge 0, L_{(p,q)} \in \naturals.
\end{equation}
\end{example}

\peter{
In the paper \cite{KriWoz19a}, the authors derived conditions equivalent to exponential strong polynomial tractability and exponential polynomial tractability, respectively. These conditions read (with notation adapted to our present paper) as follows:
\begin{itemize}
    \item We have exponential strong polynomial tractability iff there exists a $\tau>0$ and a 
    a $C\in\naturals$ such that
    \begin{equation}\label{eq:KW_exp_spt}
        \sup_{d\in\naturals} \sum_{i=C}^{\infty} \lambda_{i,d}^{i^{-\tau}} < \infty. 
    \end{equation}
    \item We have exponential polynomial tractability iff there exist $\tau_1,\tau_3\ge 0$ and $\tau_2, C>0$ such that
    \begin{equation}\label{eq:KW_exp_pt}
        \sup_{d\in\naturals} d^{-\tau_1} \sum_{i=\lceil C d^{\tau_3} \rceil}^{\infty} \lambda_{i,d}^{i^{-\tau}} < \infty. 
    \end{equation}
\end{itemize}
It should be noted that, e.g., the condition \eqref{eq:exp_spt} is easier to check in practice than the condition \eqref{eq:KW_exp_spt}. Indeed, since in \eqref{eq:KW_exp_spt} the summation indices $i$ show up in the exponents of the summands, we need the exact ordering of the eigenvalues $\lambda_{i,d}$, whereas in \eqref{eq:exp_spt} we only have a summation of the values of $1/T(\lambda_{i,d}^{-1},t,\vp)$, and the exact order of the eigenvalues can be neglected.

Secondly, we would like to show that the conditions \eqref{exp_spt} and \eqref{eq:KW_exp_spt}, as well as the conditions \eqref{exp_pt} and \eqref{eq:KW_exp_pt} are indeed equivalent. We do this for the latter two conditions, i.e., for the conditions on exponential polynomial tractability. The case of exponential strong polynomial tractability is simpler and can be treated analogously. We thus show equivalence of \eqref{eq:exp_pt} and \eqref{eq:KW_exp_pt} in the following.

Let us first assume that \eqref{eq:KW_exp_pt} holds, i.e.,
\[
M:=\sup_{d\in\naturals} d^{-\tau_1} \sum_{i=\lceil C d^{\tau_3} \rceil}^{\infty} \lambda_{i,d}^{i^{-\tau}} < \infty. 
\]
Define
\[
B_d:=\left\{i\in\naturals \colon i\ge \left\lceil Cd^{\tau_3}\right\rceil\quad\mbox{and}\quad 
\lambda_{i,d}^{i^{-\tau_2}}>\frac{1}{e}\right\}
\]
Due to our assumption, we see that $\abs{B_d} < M\,e\,d^{\tau_1}$ and $\abs{B_d}\le \lfloor M\,e\,d^{\tau_1} \rfloor$. Suppose now that $i\ge \lceil C d^{\tau_3}\rceil$ but $i\not\in B_d$, which means that 
\[\lambda_{i,d}^{i^{-\tau_2}} \le \frac{1}{e} \Leftrightarrow \log (\lambda_{i,d}^{-1})\ge i^{\tau_2}.\]
Now, choose $p> 1/\tau_2$, $q=\max\{\tau_1,\tau_3\}$, and $L_{(p,q)}=\max\{C, M\, e\}$, which implies that $\lceil L_{(p,q)} d^{q}\rceil \ge M\, e\, d^{\tau_1}$. Then, for any $d\in\naturals$, 
\begin{eqnarray*}
d^{-q} \sum_{i=\lceil L_{(p,q)} d^{q}\rceil}^\infty \frac{1}{[\max\{1,\log(1+\lambda_{i,d}^{-1})\}]^p}&=&d^{-q} \sum_{\substack{i=\lceil L_{(p,q)} d^{q}\rceil\\ i\in B_d}}^\infty \frac{1}{[\max\{1,\log(1+\lambda_{i,d}^{-1})\}]^p}\\
&& + d^{-q} \sum_{\substack{i=\lceil L_{(p,q)} d^{q}\rceil\\ i\not\in B_d}}^\infty \frac{1}{[\max\{1,\log(1+\lambda_{i,d}^{-1})\}]^p}\\
&\le& d^{-q} M\, e\, d^{\tau_1} + d^{-q}\sum_{\substack{i=\lceil L_{(p,q)} d^{q}\rceil\\ i\not\in B_d}}^\infty \frac{1}{[\log (\lambda_{i,d}^{-1})]^p}\\
&\le& 1 + d^{-q}\sum_{\substack{i=\lceil L_{(p,q)} d^{q}\rceil\\ i\not\in B_d}}^\infty \frac{1}{i^{\tau_2 p}}\\
&\le & 1 + \zeta (\tau_2 p).
\end{eqnarray*}
Since $p$ was chosen to be strictly larger than $1/\tau_2$, it follows that the latter expression is finite and its value is independent of $d$. Therefore, Condition \eqref{eq:exp_pt} holds.

Conversely, let us now assume that Condition \eqref{eq:exp_pt} holds. We will show next that this implies Condition \eqref{eq:KW_exp_pt}. Indeed, due to \eqref{eq:exp_pt}, there exists a positive constant $M$ (w.l.o.g., we can assume that $M\ge 1$) such that for all $d\in\naturals$,
\[
d^{-q} 
\sum_{i=\lceil L_{(p,q)} d^{q}\rceil}^\infty \frac{1}{[\max\{1,\log(1+\lambda_{i,d}^{-1})\}]^p} \le M,
\]
which is equivalent to 
\[
\sum_{i=\lceil L_{(p,q)} d^{q}\rceil}^\infty \frac{1}{[\max\{1,\log(1+\lambda_{i,d}^{-1})\}]^p} \le d^q M. 
\]
Similarly to the proof of Theorem \ref{thm_main_strong_tract2}, we see that for any $n > \lceil L_{(p,q)} d^{q}\rceil $, 
\[
\frac{1}{T(\lambda_{n+1,d}^{-1},1,\vp)} \le \frac{1}{n-\lceil L_{(p,q)} d^{q}\rceil +1 }
\sum_{i=\lceil L_{(p,q)} d^{q}\rceil}^\infty  \frac{1}{T(\lambda_{i,d}^{-1},1,\vp)}
\le \frac{d^q M}{n-\lceil L_{(p,q)} d^{q}\rceil +1 }
\]
This implies that
\[
\frac{n}{d^q M}-\frac{L_{(p,q)}}{M}\le\frac{n-\lceil L_{(p,q)} d^{q}\rceil +1 }{d^q M} \le [\max\{1,\log(1+\lambda_{i,d}^{-1})\}]^p 
\le [1 + \log(1+\lambda_{i,d}^{-1})]^p.
\]
Since both $L_{(p,q)}$ and $M$ in the latter chain of inequalities are independent of $d$ and $n$, there 
exists a positive constant $K$ such that 
\[
\frac{n}{d^q K} \le [1 + \log(1+\lambda_{n,d}^{-1})]^p,
\]
which implies 
\[
\frac{n^{1/p}}{d^{q/p} K^{1/p}} -1 \le \log(1+\lambda_{n,d}^{-1}).
\]
This, in turn, implies the existence of another constant $\widetilde{K}>0$ such that 
\[
\lambda_{n,d} \le \exp\left(-\frac{n^{1/p}}{\widetilde{K}^{1/p} d^{q/p}}\right)
\quad \forall n > \lceil L_{(p,q)} d^{q}\rceil,
\]
which is essentially the situation described in \cite[p.~118]{KriWoz19a}. We can then choose $\tau_2\in (0,1/p)$, 
and proceed as in \cite{KriWoz19a} to obtain 
\[
\sum_{i=\lceil L_{(p,q)} d^{q}\rceil}^\infty \lambda_{i,d}^{i^{-\tau_2}}=\mathcal{O}(d^{q/(1-\tau_2 p)}),
\]
with the implied factor in the $\mathcal{O}$-notation independent of $d$. Then, we can choose $\tau_1=q/(1-\tau_2 p)$ and $\tau_3=q$ and see that \eqref{eq:KW_exp_pt} holds.

This concludes the proof of the equivalence of \eqref{eq:exp_pt} and \eqref{eq:KW_exp_pt}.

[PLEASE CHECK CAREFULLY.]}

\kachi{In the following example, we provide a brief summary of these notions of tractability. We would like to emphasize that this example was previously known (see \cite{?}) where in this case, the roles of $\varepsilon^{-1}$ and $d$ are separated. In order to accomplish this, we will first define two functions $\phi$ and $\psi$ with positive values that are increasing in both of their arguments. This will allow us to take into account a variety of possible exponents of $\varepsilon^{-1}$ and $d$.
%To do this, we introduce two positive-value functions that are increasing in both arguments so as to take possibly different exponents of $\varepsilon^{-1}$ and $d$.
The equivalent conditions for this example are assumed to take the same form as those in Theorems \ref{thm_main_strong_tract2} and \ref{thm_main_tract2}.}

\peter{[PETER: Kachi, can you work a bit more on this introductory text, please? Right now, it does not contain much information.]}
\kachi{[KACHI: Done. Although I am having trouble adding references, do you have any suggestions.]}

\begin{example}[(Strong) Separable tractability]
Suppose that the tractability function is defined by
\[
 T(\varepsilon^{-1},d,\vp)= \phi(\varepsilon^{-1},p)\, \psi(d,q)
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, \vp \in [0,\infty)^2,
\]
with non-decreasing functions $\phi : (0,\infty) \times (0,\infty) \to (0,\infty)$ and $\psi : \naturals  \times (0,\infty) \to (0,\infty)$ and where $\vp = (p,q)$, hence $T$ is non-decreasing in all variables.
